---
title: "consultations package example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{consultations-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(consultations)
data(dummy_response, package = "consultations")
```

This package comes with some generated dummy data. This contains a few categorical columns, one column representing a multi-choice tickbox question, and one column with randomly generated free text.

```{r dummy data}
knitr::kable(dummy_response[1:5,])
```

# Categorical analysis

Any questions that are multiple choice, and/or don't have too many free-text responses, can be described and visualised in frequency tables, bar plots, and short text. This package provides functions to do this quickly. 

First, the question_types() function can detect from some rules of thumb which columns are which question type. This function returns a list with three elements: categorical, multichoice, and freetext. Each element contains a vector of column names, containing the column names which fall into each group. 

```{r qtypes}
survey_question_types(dummy_response)
```

Second, based on these question types we can calculate appropriate frequency tables. For this, we can use the response_tables() function. This returns a list, with an element for each column (ie. an element for each question). This function, by default, aggregates any answers that have fewer than 5 responses to avoid statistical disclosure: you can set this yourself using the min_n parameter.

```{r response tables}
survey_response_tables(dummy_response, survey_question_types(dummy_response), min_n = 5)
```

Third, we can describe these frequency tables as a bar plot. The survey_response_barplot() function generates a list of ggplot objects, one for each question. This is indiscriminate about question type, so for free text questions it will simply return a "Free text" bar. 

This function can take the data directly, since it runs response_tables() under the hood. 

```{r barplots}
survey_response_barplot(dummy_response)
```

Finally, we can describe this same data using text bullet points: how many possible responses were there, what was the most common, and what was the least common. The summary_text() function takes the output from response_tables, but only one element at a time, so you have to run it as part of an lapply call. 

```{r summary text}
lapply(survey_response_tables(dummy_response, qtypes = survey_question_types(dummy_response)), survey_summary_text)
```

